A SAT-Based Counterexample Guided Method for Unbounded Synthesis

### 2016-05-05 paper.tex
More minor edits

We evaluate our method on the benchmarks of the 2015 synthesis competition
(SYNTCOMP'15).  While our solver solves fewer total instances than competitors,
it solves the largest number of unique instances, i.e., instances that could not be solved by 
any other solver.  These results confirm that there exist classes of problems that are
hard for traditional synthesis techniques, but can be efficiently solved by our
method.  While further performance improvements are clearly needs, in its
current state the method may be a worthwhile addition to a portfolio solver.
%%%While further performance improvements are clearly needed, in its current state the method 
%%%can be used in a portfolio solver to increase the number of tractable instances.

@@

    One option is using a BDD solver, working backwards from the error set, to
    find all losing states. One iteration of this operation on our example
    would give the set: $\texttt{nrequests} = 3 \lor (\texttt{nrequests} = 2
    \land ( \texttt{resource0} = 0 \lor \texttt{resource1} = 0)) \lor
    (\texttt{nrequests} = 1 \land (\texttt{resource0} = 0 \land
    \texttt{resource1} = 0))$.  In the general case there is no compact
    representation of the losing set, so we try to avoid computing it by
    employing interpolation instead. The potential benefit of interpolation is
    that an interpolant can approximate the losing states in such a way that
    there is just enough information to prove realisability. It is also cheap
    to compute an interpolant using a SAT solver.

\end{exmpInt}

@@

We have discovered a set $\mathcal{I}$ of states losing for the environment.
Environment-losing states are only losing for a particular bound: given that
there does not exist an environment strategy that forces the game into an error
state in $k$ rounds or less; there may still exist a longer environment-winning
strategy.  We therefore record learned environment-losing states along with
associated bounds.  To this end, we maintain a conceptually infinite array of
sets $B^m[k]$ that are may-losing for the controller, indexed by bound $k$.
$B^m[k]$ are initialised to $\top$ for all $k$.  Whenever an environment-losing
set $\mathcal{I}$ is discovered for a node $n$ with bound $\textsc{height}(k,
n)$ in line~13 of Algorithm~\ref{alg:learn}, this set is subtracted from
$B^m[i]$, for all $i$ less than or equal to the bound (lines~14--16).

@@

The implementation of our algorithm uses \textsc{Glucose}~\cite{audemard2014}
for SAT solving and \textsc{Periplo}~\cite{rollini2013} for interpolant
generation. The benchmarks were run on a cluster of Intel Quad Core Xeon E5405
2GHz CPUs with 16GB of memory.  The solver was allowed exclusive access to a node
for one hour to solve an instance.

The results of our benchmarking are shown, along with the synthesis competition
results~\cite{syntcompedacc}, in Table~\ref{tab:syntcomp}. The competition was run on Intel Quad
Core 3.2Ghz CPUs with 32GB of memory, also on isolated nodes for one hour per
instance.

\begin{table}
    \centering
    \setlength{\tabcolsep}{8pt}
    \begin{tabular}{l | r | r }
        \textbf{Solver} & \textbf{Solved} & \textbf{Unique} \\
        \hline
        Simple BDD Solver (2) & 195 & 10 \\
        AbsSynthe (seq2) & 187 & 2 \\
        Simple BDD Solver (1) & 185 & \\
        AbsSynthe (seq3) & 179 & \\
        Realizer (sequential) & 179 & \\
        AbsSynthe (seq1) & 173 & 1 \\
        Demiurge (D1real) & 139 & 5 \\
        Aisy & 98 & \\
        \textit{Unbounded Synthesis} & \textit{103} & \textit{12} \\
    \end{tabular}
    \caption{Synthesis Competition 2015 Results}
    \label{tab:syntcomp}
\end{table}

@@

The original bounded synthesis algorithm of Narodytska et
al.~\cite{narodytska2014} solves realisability without constructing a strategy.
In \cite{een2015} the realisability algorithm is extended with strategy
extraction. The technique relies on interpolation over abstract game trees to
compute the winning strategy.  In
the present work we use interpolation in a different way in order to learn
losing states of the game.

>>>>

We evaluate our method on the benchmarks of the 2015 synthesis competition
(SYNTCOMP'15).  While our solver solves fewer total instances than competitors,
it solves the largest number of unique instances for a sequential solver, i.e.,
instances that could not be solved by any other solver.  These results confirm
that there exist classes of problems that are hard for traditional synthesis
techniques, but can be efficiently solved by our method.  While further
performance improvements are clearly needed, in its current state the method
may be a worthwhile addition to a portfolio solver.
%%%While further performance improvements are clearly needed, in its current state the method 
%%%can be used in a portfolio solver to increase the number of tractable instances.

@@

    One option is using a BDD solver, working backwards from the error set, to
    find all losing states. One iteration of this operation on our example
    would give the set: $\texttt{nrequests} = 3 \lor (\texttt{nrequests} = 2
    \land ( \texttt{resource0} = 0 \lor \texttt{resource1} = 0)) \lor
    (\texttt{nrequests} = 1 \land (\texttt{resource0} = 0 \land
    \texttt{resource1} = 0))$.  In the general case there is no compact
    representation of the losing set, so we try to avoid computing it by
    employing interpolation instead. The potential benefit of interpolation is
    that an interpolant can approximate the losing states in such a way that
    there is just enough information to prove realisability. It is also cheap
    to compute an interpolant using a SAT solver.
\end{exmpInt}

@@

\begin {exmp}

We have discovered a set $\mathcal{I}$ of states losing for the environment.
Environment-losing states are only losing for a particular bound: given that
there does not exist an environment strategy that forces the game into an error
state in $k$ rounds or less; there may still exist a longer environment-winning
strategy.  We therefore record learned environment-losing states along with
associated bounds.  To this end, we maintain a conceptually infinite array of
sets $B^m[k]$ that are may-losing for the controller, indexed by bound $k$.
$B^m[k]$ are initialised to $E$ for all $k$.  Whenever an environment-losing
set $\mathcal{I}$ is discovered for a node $n$ with bound $\textsc{height}(k,
n)$ in line~13 of Algorithm~\ref{alg:learn}, this set is subtracted from
$B^m[i]$, for all $i$ less than or equal to the bound (lines~14--16).

@@

The implementation of our algorithm uses \textsc{Glucose}~\cite{audemard2014}
for SAT solving and \textsc{Periplo}~\cite{rollini2013} for interpolant
generation. We intend to open source the tool for SYNTCOMP'16. The benchmarks
were run on a cluster of Intel Quad Core Xeon E5405 2GHz CPUs with 16GB of
memory.  The solver was allowed exclusive access to a node for one hour to
solve an instance.

The results of our benchmarking are shown, along with the synthesis competition
results~\cite{syntcompedacc}, in Table~\ref{tab:syntcomp}. The competition was
run on Intel Quad Core 3.2Ghz CPUs with 32GB of memory, also on isolated nodes
for one hour per instance. The unique column lists the number of instances that
only that tool could solve in the competition (excluding our solver). In
brackets is the number of instances that only that tool could solve, including
our solver.

\begin{table}
    \centering
    \setlength{\tabcolsep}{8pt}
    \begin{tabular}{l | r | r }
        \textbf{Solver} & \textbf{Solved} & \textbf{Unique} \\
        \hline
        Simple BDD Solver (2) & 195 & 10 \textit{(6)} \\
        AbsSynthe (seq2) & 187 & 2 \\
        Simple BDD Solver (1) & 185 & \\
        AbsSynthe (seq3) & 179 & \\
        Realizer (sequential) & 179 & \\
        AbsSynthe (seq1) & 173 & 1 \\
        Demiurge (D1real) & 139 & 5 \textit{(2)} \\
        Aisy & 98 & \\
        \textit{Unbounded Synthesis} & \textit{103} & \textit{12} \\
    \end{tabular}
    \caption{Synthesis Competition 2015 Results}
    \label{tab:syntcomp}
\end{table}

@@

The original bounded synthesis algorithm of Narodytska et
al.~\cite{narodytska2014} solves realisability without constructing a strategy.
In \cite{een2015} the realisability algorithm is extended with strategy
extraction. The technique relies on interpolation over abstract game trees to
compute the winning strategy.  In the present work we use interpolation in a
different way in order to learn losing states of the game. In addition, this
method could be easily adapted to the unbounded realisability algorithm
presented here to generate unbounded strategies.
